---
title: "Lab 05 - La Quinta is Spanish for next to Denny's, Pt. 2"
author: "Sebastian Kaltenbach"
date: "2021-10-17"
output: html_notebook
---

### Load packages and data

For our analysis, in addition to the dsbox package, we will use the following packages for data wrangling and visualisation.

*tidyverse* for data wrangling and visualization

*stringr* for fast, correct implementations of common string manipulations

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
library(stringr)
```

## Data prep

For accessability purposes, the required dataframes of Denny's and La Quinta are saved as local variables.

```{r}
dn <- dennys
lq <- laquinta
```

Load U.S. state information in local variable

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

Filter the Denny’s dataframe for Alaska (AK) and save the result as dn_ak. How many Denny’s locations are there in Alaska?
```{r}
dn_ak <- dennys %>%
  filter(state == "AK")
nrow(dn_ak)
```

### Exercise 2

Filter the La Quinta dataframe for Alaska (AK) and save the result as lq_ak. How many La Quinta locations are there in Alaska?
```{r}
lq_ak <- laquinta %>%
  filter(state == "AK")
nrow(lq_ak)
```

### Exercise 3

How many pairings are there between all Denny’s and all La Quinta locations in Alaska, i.e. how many distances do we need to calculate between the locations of these establishments in Alaska?
```{r}
nrow(dn_ak) * nrow(lq_ak)
```
Combine Denny's and La Quinta dataset for distance calculation.

```{r}
dn_lq_ak <- full_join(dn_ak, lq_ak, by = "state")
dn_lq_ak
```

### Exercise 4

How many observations are in the joined dn_lq_ak data frame? What are the names of the variables in this data frame.
```{r}
glimpse(dn_lq_ak)
```

### Exercise 5

What function from the tidyverse do we use the add a new variable to a data frame while keeping the existing variables?

The 'mutate' function.

The following function is provided by Data Science in a box and is calculating the distance between two locations (lat, long).

```{r}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 = long1 * pi / 180
  lat1  = lat1  * pi / 180
  long2 = long2 * pi / 180
  lat2  = lat2  * pi / 180
  
  R = 6371 # Earth mean radius in km
  
  a = sin((lat2 - lat1)/2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1)/2)^2
  d = R * 2 * asin(sqrt(a))
  
  return( round(d,round) ) # distance in km
}
```

### Exercise 6

Calculate the distances between all pairs of Denny’s and La Quinta locations and save this variable as distance. Make sure to save this variable in THE dn_lq_ak data frame so that you can use it later.
```{r}
dn_lq_ak <- dn_lq_ak %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y))
dn_lq_ak
```

### Exercise 7

Calculate the minimum distance between a Denny’s and La Quinta for each Denny’s location. To do so we group by Denny’s locations and calculate a new variable that stores the information for the minimum distance.

```{r}
dn_lq_ak_mindist <- dn_lq_ak %>%
  group_by(address.x) %>%
  summarise(closest = min(distance))
dn_lq_ak_mindist
```

### Exercise 8

Describe the distribution of the distances Denny’s and the nearest La Quinta locations in Alaska. Also include an appripriate visualization and relevant summary statistics.

```{r}
ggplot(data = dn_lq_ak_mindist, aes(y = closest)) +
  geom_boxplot(fill="cadetblue1") + 
  theme_bw() + 
  labs(title = "Distance distribution of Denny's and La Quinta in Alaska I", 
       x = "", y = "Distance [km]")

ggplot(data = dn_lq_ak_mindist, aes(x = closest)) + 
 geom_dotplot(binwidth = 0.25) +
  theme_bw() + 
  labs(title = "Distance distribution of Denny's and La Quinta in Alaska II", 
       x = "Distance [km]", y = "")
```

### Exercise 9

Repeat the same analysis for North Carolina: (i) filter Denny’s and La Quinta Data Frames for NC, (ii) join these data frames to get a completelist of all possible pairings, (iii) calculate the distances between all possible pairings of Denny’s and La Quinta in NC, (iv) find the minimum distance between each Denny’s and La Quinta location, (v) visualize and describe the distribution of these shortest distances using appropriate summary statistics.

```{r}
dn_lq_nc <- full_join(dennys %>% filter(state == "NC"), laquinta %>% filter(state == "NC"), by = "state") %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y))
dn_lq_nc
```
```{r}
dn_lq_nc_mindist <- dn_lq_nc %>%
  group_by(address.x) %>%
  summarise(closest = min(distance))
dn_lq_nc_mindist
```
```{r}
ggplot(data = dn_lq_nc_mindist, aes(y = closest)) +
  geom_boxplot() +
  geom_boxplot(fill="cadetblue1") + 
  theme_bw() +
  labs(title = "Distance distribution of Denny's and La Quinta in North Carolina I", 
       x = "", y = "Distance [km]")

ggplot(data = dn_lq_nc_mindist, aes(x = closest)) + 
 geom_dotplot(binwidth = 10) +
  theme_bw() + 
  labs(title = "Distance distribution of Denny's and La Quinta in North Carolina II", 
       x = "Distance [km]", y = "")
```

### Exercise 10

Repeat the same analysis for Texas.

```{r}
dn_lq_tx <- full_join(dennys %>% filter(state == "TX"), laquinta %>% filter(state == "TX"), by = "state") %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y))
dn_lq_tx
```
```{r}
dn_lq_tx_mindist <- dn_lq_tx %>%
  group_by(address.x) %>%
  summarise(closest = min(distance))
dn_lq_tx_mindist
```
```{r}
ggplot(data = dn_lq_tx_mindist, aes(y = closest)) +
  geom_boxplot() + 
  geom_boxplot(fill="cadetblue1") + 
  theme_bw() +
  labs(title = "Distance distribution of Denny's and La Quinta in Texas I", 
       x = "", y = "Distance [km]")

ggplot(data = dn_lq_tx_mindist, aes(x = closest)) + 
 geom_dotplot(binwidth = 0.65) +
  theme_bw() + 
  labs(title = "Distance distribution of Denny's and La Quinta in Texas II", 
       x = "Distance [km]", y = "")
```

### Exercise 11

Repeat the same analysis for a state of your choosing, different than the ones we covered so far.

State New York - NY

```{r}
dn_lq_ny <- full_join(dennys %>% filter(state == "NY"), laquinta %>% filter(state == "NY"), by = "state") %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y))
dn_lq_ny
```
```{r}
dn_lq_ny_mindist <- dn_lq_ny %>%
  group_by(address.x) %>%
  summarise(closest = min(distance))
dn_lq_ny_mindist
```
```{r}
ggplot(data = dn_lq_ny_mindist, aes(y = closest)) +
  geom_boxplot() + 
  geom_boxplot(fill="cadetblue1") + 
  theme_bw() +
  labs(title = "Distance distribution of Denny's and La Quinta in New York I", 
       x = "", y = "Distance [km]")

ggplot(data = dn_lq_ny_mindist, aes(x = closest)) + 
 geom_dotplot(binwidth = 5) +
  theme_bw() + 
  labs(title = "Distance distribution of Denny's and La Quinta in New York II", 
       x = "Distance [km]", y = "")
```

### Exercise 12

Among the states you examined, where is Mitch Hedberg’s joke most likely to hold true? Explain your reasoning.

Most likely the joke holds in Alaska, North Carolina, and Texas. The distribution of the closest distance are in these areas around 5 km (median) which could be defined as 'close'. Texas should be mentioned, that the distribution is very broad and therefore not meaningful for this study (Not every La Quinta location is close to a Denny's). 

Besides, New York as self-defined state in the U.S. does not hold the joke, because the median distance between both Denny's and La Quinta is 25 km, which is not close compared to the other states.


## Function - Create dataframe by state
1. Get all of lab 5 working, to include creating a function that takes a "state" as input and produces the output dataframe of distinces.
2. Modify (probably already done in class) so that ... given a state, if creates a dataframe that for each la quinta, it includes distance to nearest dennys.

```{r}
analyze_state <- function(this_state) {
  
  # Define Denny's dataframe filtered by state
  dn_state <- dn %>%
    filter(state == this_state)
  # Define La Quinta's dataframe filtered by state  
  lq_state <- lq %>%
    filter(state == this_state)
  # Define dataframe with min distances and calculate distances between two locations with the haversine function
  dm_lq_state_mindist <-- full_join(dn_state, lq_state, by = "state") %>%
    mutate(
      distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 4)
    ) %>%
    group_by(address.x) %>%
    summarise(closest = min(distance))
  # Return the final dataframe for further processing
  return( dm_lq_state_mindist )
}
```


3. Grab the dataset (available in moodle in "week of 28 Sep" ) for the 10000 US restaurants.

```{r}
FastFoodRestaurants <- read_csv("FastFoodRestaurants.csv")
# Clean the existing FastFoodRestaurants dataframe
FastFoodRestaurants <- FastFoodRestaurants %>% 
  rename(state = province) %>%
  mutate(name = tolower(name),
         name = str_replace_all(name, fixed(" "), ""),
         name = str_replace_all(name, fixed("'"), ""),
         name = str_replace_all(name, fixed("/"), ""),
         name = str_replace_all(name, fixed("-"), ""))
```

4. Create a function that takes a "state" as input and produces a dataframe, where there is one observation per "la quinta" and it lists distance to nearest "Restaurant". WHEN YOU BUILD THIS FUNCTION ... ensure it is pulling a sample from the state of interest AND that the sample size is similar to the number of Denny's in the state.  (Hint - use "sample_n" function from dplyr.)

```{r}
analyze_distance_with_selection <- function(this_state, this_restaurant) {
  # Define Denny's dataframe filtered by state  
  dn_state <- dn %>%
    filter(state == this_state)
  # Define La Quinta's dataframe filtered by state   
  lq_state <- lq %>%
    filter(state == this_state)
  # Define FastFoodRestaurants's dataframe filtered by state and specified fast food chain  
  fastfood_state <- FastFoodRestaurants %>%
    filter(state == this_state) %>%
    filter(name == this_restaurant)
  # create fastfood dataframe sample according to size of denny's dataframe (for comparison reasons)
  fastfood_sample <- sample_n(fastfood_state, nrow(dn_state))
  # Define dataframe with min distances and calculate distances between two locations with the haversine function - La Quinta and Denny's
  dn_mindist <- full_join(lq_state, dn_state, by = "state") %>%
    mutate(
      distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 4)
    ) %>%
    group_by(address.x) %>%
    summarise(closest = min(distance)) %>%
    mutate(name = "dennys")
  fastfood_sample <- sample_n(fastfood_state, nrow(dn_state))
  # Define dataframe with min distances and calculate distances between two locations with the haversine function - La Quinta and specific fast food restaurant
  ff_mindist <- full_join(lq_state, fastfood_sample, by = "state") %>%
    mutate(
      distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 4)
    ) %>%
    group_by(address.x) %>%
    summarise(closest = min(distance)) %>%
    mutate(name = this_restaurant)
  # Define final dataframe with all required data
  df_final <- full_join(dn_mindist, ff_mindist, by = "address.x")
  # Return the final dataframe for further processing  
  return (df_final)
}
```

5. Do some comparative analysis on several states of interest.   Are "denny's" any closer than the "other" restaurant?

```{r}

df_lq_dn_mc_ak <- analyze_distance_with_selection("AK", "mcdonalds")

ggplot(data = df_lq_dn_mc_ak, aes(y = closest.x)) +
  geom_boxplot() + 
  geom_boxplot(fill="cadetblue1") + 
  theme_bw() +
  labs(title = "Distance distribution of Denny's and La Quinta in Alaska", 
       x = "", y = "Distance [km]")
ggplot(data = df_lq_dn_mc_ak, aes(y = closest.y)) +
  geom_boxplot() + 
  geom_boxplot(fill="orange1") + 
  theme_bw() +
  labs(title = "Distance distribution of McDonald's and La Quinta in Alaska", 
       x = "", y = "Distance [km]")


df_lq_dn_mc_ny <- analyze_distance_with_selection("NY", "mcdonalds")

ggplot(data = df_lq_dn_mc_ny, aes(y = closest.x)) +
  geom_boxplot() + 
  geom_boxplot(fill="cadetblue1") + 
  theme_bw() +
  labs(title = "Distance distribution of Denny's and La Quinta in New York", 
       x = "", y = "Distance [km]")
ggplot(data = df_lq_dn_mc_ny, aes(y = closest.y)) +
  geom_boxplot() + 
  geom_boxplot(fill="orange1") + 
  theme_bw() +
  labs(title = "Distance distribution of McDonald's and La Quinta in New York", 
       x = "", y = "Distance [km]")

df_lq_dn_mc_nc <- analyze_distance_with_selection("NC", "mcdonalds")

ggplot(data = df_lq_dn_mc_nc, aes(y = closest.x)) +
  geom_boxplot() + 
  geom_boxplot(fill="cadetblue1") + 
  theme_bw() +
  labs(title = "Distance distribution of Denny's and La Quinta in North Carolina", 
       x = "", y = "Distance [km]")
ggplot(data = df_lq_dn_mc_nc, aes(y = closest.y)) +
  geom_boxplot() + 
  geom_boxplot(fill="orange1") + 
  theme_bw() +
  labs(title = "Distance distribution of McDonald's and La Quinta in North Carolina", 
       x = "", y = "Distance [km]")

df_lq_dn_mc_ma <- analyze_distance_with_selection("MA", "mcdonalds")

ggplot(data = df_lq_dn_mc_ma, aes(y = closest.x)) +
  geom_boxplot() + 
  geom_boxplot(fill="cadetblue1") + 
  theme_bw() +
  labs(title = "Distance distribution of Denny's and La Quinta in Massachusetts", 
       x = "", y = "Distance [km]")
ggplot(data = df_lq_dn_mc_ma, aes(y = closest.y)) +
  geom_boxplot() + 
  geom_boxplot(fill="orange1") + 
  theme_bw() +
  labs(title = "Distance distribution of McDonald's and La Quinta in Massachusetts", 
       x = "", y = "Distance [km]")
```
## Are "denny's" any closer than the "other" restaurant?

The "other" restaurant we used for comparison was McDonald's. Denny's were closer to LaQuinta in some of the states we analyzed, but not all. We don't know the actual answer to the question, since FastFoodRestaurants is a sample, and we also took a random sample of McDonald's in the states we chose. This means that there may be closer McDonald's to LaQuinta's that are not in our sample. However, we did not sample the Denny's and LaQuinta locations. 