---
title: "Lab4DSbox"
author: "Tarak Patel"
date: "10/8/2021"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load packages and data

```{r load-packages, message = FALSE}
#install.packages("devtools")
#devtools::install_github("rstudio-education/dsbox")

library(tidyverse) 
library(dsbox) 
library(readxl)
library(here)

```
### Loading state data
```{r load-data, message = FALSE}
states <- read_excel(here("data","raw","states.xlsx"))
```

### Exercise 1

What are the dimensions of the Denny’s dataset? (Hint: Use inline R code and functions like nrow and ncol to compose your answer.) What does each row in the dataset represent? What are the variables?

```{r}
nrow(dennys)
ncol(dennys)
```

Rows: 1643
Columns: 6
In the dataset, each row represents a different Denny's location.
The variables are: address, city, state, zip, longitude, latitude

### Exercise 2

What are the dimensions of the La Quinta’s dataset? What does each row in the dataset represent? What are the variables?

```{r}
nrow(laquinta)
ncol(laquinta)
```
Rows: 909
Columns: 6
Each row represents a different La Quinta location.
The variable are: address, city, state, zip, longitude, latitude 

### Exercise 3

Take a look at the websites that the data come from (linked above). Are there any La Quinta’s locations outside of the US? If so, which countries? What about Denny’s?

There are no Denny's outside of the US. There are international La Quinta locations. There are locations in Canada, Mexico, New Zealand, Honduras, Turkey, UAE, Chile, and Columbia

### Exercise 4

Now take a look at the data. What would be some ways of determining whether or not either establishment has any locations outside the US using just the data (and not the websites). Don’t worry about whether you know how to implement this, just brainstorm some ideas. Write down at least one as your answer, but you’re welcomed to write down a few options too.

You could filter by state to see if there are results that are not a US state.

### Exercise 5

Find the Denny’s locations that are outside the US, if any. To do so, filter the Denny’s locations for observations where state is not in states$abbreviation. The code for this is given below. Note that the %in% operator matches the states listed in the state variable to those listed in states$abbreviation. The ! operator means not. Are there any Denny’s locations outside the US?

```{r}
dennys %>%
  filter(!(state %in% states$abbreviation))
```
There are no Denny's locations outside the US.

### Exercise 6

Add a country variable to the Denny’s dataset and set all observations equal to "United States". Remember, you can use the mutate function for adding a variable. Make sure to save the result of this as dennys again so that the stored data frame contains the new variable going forward.

```{r}
dennys %>%
  mutate(country = "United States")
```


### Exercise 7

Find the La Quinta locations that are outside the US, and figure out which country they are in. This might require some googling. Take notes, you will need to use this information in the next exercise.

Countries are listed above in Exercise 3.

### Exercise 8

Add a country variable to the La Quinta dataset. Use the case_when function to populate this variable. You’ll need to refer to your notes from Exercise 7 about which country the non-US locations are in.

```{r}
laquinta %>%
  mutate(country = case_when(
    state %in% state.abb     ~ "United States",
    state %in% c("ON", "BC") ~ "Canada",
    state == "ANT"           ~ "Colombia",
    state == "FM"            ~ "Honduras",
    state %in% c("AG","CH","NL", "PU", "QR", "SL", "VE") ~ "Mexico"))

```

### Exercise 9
```{r}
new_laquinta <- laquinta %>%
  filter(country == "United States")
```

Which states have the most and fewest Denny’s locations? What about La Quinta? Is this surprising? Why or why not?

```{r}
dennys %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation"))

```
```{r}
laquinta %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation"))
```

Denny's 3 Most: California, Florida, Texas
Denny's 3 Least: Delaware, DC, Alaska, Vermont
La Quinta 3 Most: Texas, Florida, California, 
La Quinta 3 Least: Alaska, Maine, Vermont

It is not surprising that the larger states have more locations.

### Exercise 10

Which states have the most Denny’s locations per thousand square miles? What about La Quinta?

```{r}
dennys <- dennys %>%
  mutate(establishment = "Denny's")
laquinta <- laquinta %>%
  mutate(establishment = "La Quinta")
dn_lq <- bind_rows(dennys, laquinta)
ggplot(dn_lq, mapping = aes(x = longitude, y = latitude, color = establishment)) +
  geom_point()
```


### Exercise 11

Filter the data for observations in North Carolina only, and recreate the plot. You should also adjust the transparency of the points, by setting the alpha level, so that it’s easier to see the overplotted ones. Visually, does Mitch Hedberg’s joke appear to hold here?

```{r}
new_dn <- dennys %>%
  filter(state == "NC")
new_lq <- laquinta %>%
  filter(state == "NC")
dn_lq_2 <- bind_rows(new_dn, new_lq)
ggplot(dn_lq_2, mapping = aes(x = longitude, y = latitude, color = establishment)) +
  geom_point()
```


### Exercise 12

Now filter the data for observations in Texas only, and recreate the plot, with an appropriate alpha level. Visually, does Mitch Hedberg’s joke appear to hold here?

```{r}
new_dn_3 <- dennys %>%
  filter(state == "TX")
new_lq_3 <- laquinta %>%
  filter(state == "TX")
dn_lq_3 <- bind_rows(new_dn_3, new_lq_3)
ggplot(dn_lq_3, mapping = aes(x = longitude, y = latitude, color = establishment)) +
  geom_point()
```


